---
title: "Camille's R Book"
author: "Camille Fairbourn"
output: html_document
---
#Module 3 Discrete Random Variables

As mentioned previously, each time you open a new RStudio session, you need to run the following three commands.

```{r message = FALSE}
require(mosaic)
require(openintro)
require(MASS)
```
####Creating a probability distribution from calculated theoretical probabilities

In the lectures we see a graph of the discrete random variable for X, the number of girls in a 3-child family

We'll use the combine function c() to create a vector of values for the random variable X and then use it again to create a vector for the probabilities.

```{r}
values<-c(0, 1, 2, 3)
probs<-c(1/8, 3/8, 3/8, 1/8)
```

Next, we'll use the barplot() function to create a barchart of the probability distribution

```{r}
barplot(probs, # Specify the vector describing the bars to plot
        names.arg = values, # Specify the vector of names to be plotted below each bar
        ylim = c(0,.4), # Specify the limits for the y-axis
        main = "3 child families", # Create a title for the chart
        ylab = "Probability", # Label the y-axis
        xlab = "Number of girls", #Label the x-axis
        col = "#D4A7B1") #Specify the color. Here I used a hexadecimal value.
```

####Creating a probability distribution from observed frequencies

Example: Restaurant Seating

Again we use the combine function, c(), to create a vector of values for the random variable X, the number of people in a restaurant party. We'll also use it to create a vector of the observed frequency for each value of X.

```{r}
values <- c(1, 2, 3, 4, 5, 6, 7)
freq <- c(25, 89, 101, 96, 58, 39, 12)
```

To calculate the probabilities, we need to divide each frequency by the total. The function sum() will calculate the total for us.

```{r}
prob <- freq / sum(freq)
```

As above, we use the barplot() function to create the graph of the *frequencies*.

```{r}
barplot(freq, # Specify the vector describing the bars to plot
        names.arg = values, # Specify the vector of names to be plotted below each bar
        main = "Restaurant Seating", # Create a title for the chart
        ylab = "Frequency", # Label the y-axis
        xlab = "Party Size", #Label the x-axis
        col = "#F3A176") #Specify the color. Here I used a hexadecimal value.
```

We can also use barplot() to create the probability graph.

```{r]}
barplot(prob, # Specify the vector describing the bars to plot
        names.arg = values, # Specify the vector of names to be plotted below each bar
        ylim = c(0,.25), # Specify the limits for the y-axis
        main = "Restaurant Seating", # Create a title for the chart
        ylab = "Probability", # Label the y-axis
        xlab = "Party Size", #Label the x-axis
        col = "#D4A7B1") #Specify the color. Here I used a hexadecimal value.
```

Notice that while the two graphs appear visually similar, in the one we are plotting frequencies and the other probabilities. This is reflected in the y-axis.