---
title: "USU STAT 2300 Module 4.2"
author: "Camille Fairbourn"
output:
      html_document:
        keep_md: true
---

#Module 4, Part 2: Sampling Distributions

##Sampling Distributions for the Mean

As mentioned previously, each time you open a new RStudio session, you need to run the following three commands.

```{r message = FALSE}
require(mosaic)
require(openintro)
require(MASS)
```

The U.S. Department of Education hosts a [website](https://collegescorecard.ed.gov/data/) where they make available data about all undergraduate degree-granting institutions of higher education in the United States. In this exercise, we will work with a sample of 506 public universities and their recorded information from the year 2013.

First, we read in the dataset and construct a density histogram of the average yearly cost of attendance. We will also calculate and plot a vertical line at the mean of this variable.
```{r}
collegecost <- read.csv("http://www.math.usu.edu/cfairbourn/Stat2300/RStudioFiles/data/collegecost.csv")

hist(collegecost$yearly,
     main = "Public Universities in the United States, 2013",
     xlab = "Average yearly cost of attendance in $",
     prob = TRUE)
abline(v = mean(collegecost$yearly), col = "red", lwd = 3)
mean(collegecost$yearly)
```

Let's take a simple random sample of 5 of these universities and store the results in a vector called x. We will also calculate the mean for the sample.

```{r}
x <- sample(collegecost$yearly, 5)
#look at the numbers in this sample and calculate the mean
x #data for this sample
mean(x) #this sample mean
```

Run the code below 3 more times and make note of the mean of each sample

```{r eval = FALSE}
x <- sample(collegecost$yearly, 5)
x
mean(x)
``` 

**Increase the sample size to n = 20**

Let's repeat this for a larger sample size. Make note of the mean of each sample.

```{r}
x <- sample(collegecost$yearly, 20)
x #data for this sample
mean(x) #this sample mean
```

Run the code below 3 more times and make note of the mean of each sample

```{r eval = FALSE}
x <- sample(collegecost$yearly, 20)
x
mean(x)
``` 

**Increase the sample size to n = 50**

Repeat once more for an even larger sample size. Again, make a note of the mean of each sample.

```{r}
x <- sample(collegecost$yearly, 50)
x #data for this sample
mean(x) #this sample mean
```

Run the code below 3 more times and make note of the mean of each sample

```{r eval = FALSE}
x <- sample(collegecost$yearly, 50)
x
mean(x)
``` 

Now we're going to have R take 500 samples of n=5 universities and record the mean of each sample.

Specify the sample size, n
```{r}
n <- 5
#Create a vector to store the sample means and draw the samples
xbar = rep(0,500)
for(i in 1:500) {xbar[i] = mean(sample(collegecost$yearly, n))}

#Calculate the mean of the 500 sample means and compare it to 
#the population mean
mean(xbar) #the mean of the sample means
mean(collegecost$yearly) #the population mean

#Create a histogram of the 500 sample means 
#with a line at the population mean
hist(xbar, 
     prob = TRUE, 
     breaks=12, 
     xlim = c(5000, 35000),
     main="Sample Means",
     xlab="Mean")
legend("topright",c("n=",n))
abline(v=mean(collegecost$yearly), col="red", lwd=2)

#Compare the sample means histogram to the data histogram. 
#What do they have in common? How are they different?

#Let's change the sample size to n=20 and repeat.
n<-20
#Create a vector to store the sample means and draw the samples
xbar=rep(0,500)
for(i in 1:500) {xbar[i]=mean(sample(collegecost$yearly, n))}

#Calculate the mean of the 500 sample means and compare it to 
#the population mean
mean(xbar) #the mean of the sample means
mean(collegecost$yearly) #the population mean

#Create a histogram of the 500 sample means 
#with a line at the population mean
hist(xbar, 
     prob = TRUE, 
     breaks=12, 
     xlim = c(5000, 35000),
     main="Sample Means",
     xlab="Mean")
legend("topright",c("n=",n))
abline(v=mean(collegecost$yearly), col="red", lwd=2)

#Now change the sample size to n=50 and repeat.
n<-50
#Create a vector to store the sample means and draw the samples
xbar=rep(0,500)
for(i in 1:500) {xbar[i]=mean(sample(collegecost$yearly, n))}

#Calculate the mean of the 500 sample means and compare it to 
#the population mean
mean(xbar) #the mean of the sample means
mean(collegecost$yearly) #the population mean

#Create a histogram of the 500 sample means 
#with a line at the population mean
hist(xbar, 
     prob = TRUE, 
     breaks=12, 
     xlim = c(5000, 35000),
     main="Sample Means",
     xlab="Mean")
legend("topright",c("n=",n))
abline(v=mean(collegecost$yearly), col="red", lwd=2)