---
title: "Camille's R Book"
author: "Camille Fairbourn"
output:
      html_document:
        keep_md: true
---

#Module 4, Part 1: Continuous Random Variables

##Continuous Random Variables

As mentioned previously, each time you open a new RStudio session, you need to run the following three commands.

```{r message = FALSE}
require(mosaic)
require(openintro)
require(MASS)
```

We're going to look at the Landfill data set, a sample of 200 weights of trucks as they enter the landfill.

First, we read in the dataset.
```{r}
Landfill<-read.csv("http://www.math.usu.edu/cfairbourn/Stat2300/RStudioFiles/data/Landfill.csv")
```

We'll use the `hist()` function to construct a frequency histogram of the truck weights, and include a vertical line at x = 38000.
```{r}
hist(Landfill$Weights, 
     main = "Truck Weights", #Include a chart title
     xlab = "Weight in pounds", #Change the label for the x-axis
     col = "lightskyblue", #Change the fill color of the histogram
     xaxt = 'n') #This leaves the x-axis blank so we can specify the tick marks ourselves.
axis(side=1, at=seq(30000,55000,2000)) #This sets the tick marks every 2000, starting at 30000 and ending at 55000.
abline(v=38000, col="red", lwd=3) #Add in a thick vertical line at x=38000
```
#Count the trucks that come in underweight
tally(~Weights<=38000, data=Landfill)

#Construct a density histogram of the truck weights
hist(Landfill$Weights, main="Truck Weights", prob=TRUE,
     xlab="Weight in pounds", col="lightskyblue", xaxt='n')
axis(side=1, at=seq(30000,55000,2000))
#Add in a thick vertical line at x=38000
abline(v=38000, col="red", lwd=3)

#Add in the normal curve overlay
points(seq(min(Landfill$Weights)-1500, max(Landfill$Weights)+1500, length.out=500),
       dnorm(seq(min(Landfill$Weights)-1500, max(Landfill$Weights)+1500, length.out=500),
             mean(Landfill$Weights), sd(Landfill$Weights)), type="l", col="darkblue", lwd=2)

#Calculate the area under the normal curve less than 38000
pnorm(38000, mean=mean(Landfill$Weights), sd=sd(Landfill$Weights))