---
title: "USU STAT 2300 Module 4.2"
author: "Camille Fairbourn"
output:
      html_document:
        keep_md: true
---

#Module 4, Part 2: Sampling Distributions

##Sampling Distributions for the Mean

As mentioned previously, each time you open a new RStudio session, you need to run the following three commands.

```{r message = FALSE}
require(mosaic)
require(openintro)
require(MASS)
```

The U.S. Department of Education hosts a [website](https://collegescorecard.ed.gov/data/) where they make available data about all undergraduate degree-granting institutions of higher education in the United States. In this exercise, we will work with a sample of 506 public universities and their recorded information from the year 2013.

First, we read in the dataset and construct a density histogram of the average yearly cost of attendance. We will also calculate and plot a vertical line at the mean of this variable.
```{r}
collegecost<-read.csv("http://www.math.usu.edu/cfairbourn/Stat2300/RStudioFiles/data/collegecost.csv")

hist(collegecost$yearly,
     main = "Public Universities in the United States, 2013",
     xlab = "Average yearly cost of attendance in $",
     prob = TRUE)
abline(v = mean(collegecost$yearly), col = "red", lwd = 3)
mean(collegecost$yearly)
```

Let's take a simple random sample of 5 of these universities and store the results in a vector called x. We will also calculate the mean for the sample.

```{r}
x<-sample(collegecost$yearly, 5)
#look at the numbers in this sample and calculate the mean
x #data for this sample
mean(x) #this sample mean



#Now take a simple random sample of 20 of these trucks and
#store the results in a vector called x.
x<-sample(collegecost$yearly, 20)
#look at the numbers in this sample and calculate the mean
x #data for this sample
mean(x) #this sample mean

#Re-run lines 34 through 37 three times and record your sample 
#mean each time on the worksheet. Mark each sample mean on 
#your histogram.

#Now take a simple random sample of 50 of these trucks and
#store the results in a vector called x.
x<-sample(collegecost$yearly, 50)
#look at the numbers in this sample and calculate the mean
x #data for this sample
mean(x) #this sample mean
```